---
title: "Merging_Databases"
author: "Oskar Nyberg"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
  library(tidyverse)
  library(readxl)
```

# Merging HESTIA Tox DB with the EnviroTox DB.
The code below adds Envirotox "on top of" the HESTIA DB, Prioritizing HESTIA CASRN and Species, while adding CAS and species' unique to the Envirotox DB.

```{r Envirotox}
# Specifying which columns i want to keep in the main dataset
common_name_elements <- c(
  "CAS.Number", "DB", "version", "Database", "Species", "Genus", 
  "Taxonomy.Group", "Medium", "Effect", "Value.mg_l", "AcuteChronic", 
  "Endpoint", "Endpoint_conv", "Time.Hours",
  "EC10eq", "EC10eq_Chronic", "EC10eq_Acute", "No.Extrapolations", "Source")

# Reading in the Envirotox database (created in ../code/Envirotox_DB.Rmd)
EnviroTox_toxDB_EC10eq <- read.csv("../data/RAW_DB/EnviroTox_toxDB_EC10eq.csv")

# filtering out the unique Envirotox records
EnviroTox_Q_dat <- rbind(
  EnviroTox_toxDB_EC10eq %>% 
    # Removing all species taxonomic annotations given as "Genus sp."
    filter(!grepl("sp\\.", Species)) %>% 
    # selecting all CAS from `EnviroTox_substDB` that are NOT overlapping.
    filter(!CAS.Number %in% Q_dat$CAS.Number),
  EnviroTox_toxDB_EC10eq %>% 
    # Removing all species taxonomic annotations given as "Genus sp."
    filter(!grepl("sp\\.", Species)) %>% 
    # selecting all species that are unique to the `Q_dat` DB- 
    filter(CAS.Number %in% Q_dat$CAS.Number,
         !Species %in% Q_dat$Species))  %>% 
  select(common_name_elements)

# Adjusting the content of the HESTIA database
HESTIA_BASE_dat <- "../results/HESTIA_EC10eq_DB.csv" %>% 
  rename(Medium = Media.type) %>% 
  unite(Source, c("Title", "Author", "Year", "Reference.source", "URL", "DOI"), sep = ";", na.rm = TRUE) %>% 
  select(common_name_elements)

# This is the raw database with HESTIA as BASE and Envirotox as "filler"
HESTIA_BASE_EnviroTox_FILL <- rbind(HESTIA_BASE_dat, EnviroTox_Q_dat)

write.csv(HESTIA_BASE_EnviroTox_FILL, "../results/FINAL_HESTIA_BASE_EnviroTox_FILL.csv", row.names = F)
write.csv(HESTIA_BASE_EnviroTox_FILL, "../HESTIA_Shiny/data/FINAL_HESTIA_BASE_EnviroTox_FILL.csv", row.names = F)
```

## Merging all physchem & subst. use-type annotations 
```{r MergingPhyschem}
FULL_PHYSCHEM <- rbind(
  NEW_PHYSCHEM,
  read.csv("../data/excel_references/Envirotox_physchem.csv") %>% 
    # Duplicates in the two datasets are present. Filtering out these here.
    filter(!CAS.Number %in% NEW_PHYSCHEM$CAS.Number)) %>% 
  mutate(KdegA = case_when(KdegA == 0 ~ as.numeric(NA), TRUE ~ KdegA))
```


## Calculating HC50_EC50 for the HESTIA_envirotox Database for backwards compatibility with previous LCA-tox versions.
Dependency on `../code/HC50_calculation_function.R`-file with the defined function. 
This operation requires Physchem/subst.use-group data to be joined with the toxicological data.
```{r HC50EC50}
source("../code/HC50_calculation_function.R")

# Attaching the Physchem data to HESTIA dataset and outputting a df with HC50 data
HESTIA_HC50_calc <- HC50_calc_function(HESTIA_BASE_EnviroTox_FILL %>% left_join(x = ., y = FULL_PHYSCHEM, by = "CAS.Number"))
```

## HC20EC10eq calculations for HESTIA_BASE_EnviroTox_FILL

```{r HC20Calculations}
source("../code/HC20_calculation_function.R")

HESTIA_HC20_dataset <- HCx_calculator(HESTIA_BASE_EnviroTox_FILL, HCx = 20)

write.csv(HESTIA_HC20_dataset, "../results/HESTIA_HC20_dataset.csv", row.names = F)
```

## Running the weighted uncertainty model
```{r}
# Dataset to analyze is object `HESTIA_BASE_EnviroTox_FILL`

# load the nls function
 source("../code/least_squares_fit_model_code.R")
 # Running the operation with the HESTIA toxicity data set, generating a list object
system.time(nls_output <- nls_across_all(dataset = HESTIA_BASE_EnviroTox_FILL, Plot_output = "NO", MC_n = 1e+05, Plot_destination = "../figures/SSD_plots"))

#Transform into a dataframe - removing the list of lists containing full nls-output
nls_output_df <- as.data.frame(do.call(cbind, nls_output[1:16])) %>%
 mutate(across(c(2:15), ~ as.numeric(.x)),
        CAS.Number = as.factor(CAS.Number),
        status = as.character(status)) #%>% 
  # mutate(status = case_when(status == "warning: step factor reduced below 'minFactor'" ~ "Not enough data",
  #                           status == "singular gradient matrix at initial parameter estimates" ~ "One species record",
  #                           TRUE ~ status))

write.csv(nls_output_df, "../results/nls_output_df.csv", row.names = F)
write.csv(nls_output_df, "../HESTIA_Shiny/data/nls_output_df.csv", row.names = F)

nls_output_df <- read.csv("../results/nls_output_df.csv")

```


## Merging Physchem & Tox data
```{r FinalDataset}
HESTIA_envirotox_cfs <- left_join(
    x = FULL_PHYSCHEM, 
    y = HESTIA_HC20_dataset,
    by = "CAS.Number") %>% 
# Selection for the SHINY-app
  select(c(
    CAS.Number, PesticideAI_name, CanonicalSMILES, 
    MW.g.mol, pKaChemClass, pKa.gain, pKa.loss, Kow_L.L, Koc_L.kg_MCI,
    kH25C_Pa.m3.mol, Vapor.Pressure_Pa, Sol_mg.L,
    KdegA, KdegW, KdegSd, KdegSl, HC20, BAF_L.Kg, 
    mean_EC10eq, Sd_EC10eq, HC20EC10eq,
    Molecular.formula, Predefined.substance.type, Substance_type, Group, Subgroup, Heavy.Metals, Halogenated, definition, n_data_recs, n_species, sum_extrapolations, n_Taxonomy.Group 
    )
  ) %>% 
  mutate(across(contains("Kdeg"), ~ formatC(., format = "e")))

write.csv(HESTIA_envirotox_cfs, "../HESTIA_Shiny/data/HESTIA_envirotox_cfs.csv", row.names = F)

```

I'm reading in the USEtox Organics database names, to be able to export an identical list of data for easier USEtox input
```{r}

USEtox_input_names <- names(read_xlsx("../data/Impact assessment/USEtox_substance_data_organics.xlsx", sheet = "Substance data", skip = 1))
```

Creating an output file with the USEtox 2.1 database input structure 
Defining a data frame shaped like the USEtox-input database

```{r}
USEtox_input <- data.frame(
 ...1 = seq(1, nrow(HESTIA_envirotox_cfs), 1),
CASRN = HESTIA_envirotox_cfs$CAS.Number,
Name = HESTIA_envirotox_cfs$PesticideAI_name,
PesticideTargetClass = HESTIA_envirotox_cfs$Group,
PesticideChemClass = HESTIA_envirotox_cfs$Substance_type,
MW = as.numeric(HESTIA_envirotox_cfs$MW.g.mol),
pKaChemClass = HESTIA_envirotox_cfs$pKaChemClass,
pKa.gain = HESTIA_envirotox_cfs$pKa.gain,
pKa.loss = HESTIA_envirotox_cfs$pKa.loss,
KOW = HESTIA_envirotox_cfs$Kow_L.L,
Koc = HESTIA_envirotox_cfs$Koc_L.kg_MCI,
KH25C = HESTIA_envirotox_cfs$kH25C_Pa.m3.mol,
Pvap25 = HESTIA_envirotox_cfs$Vapor.Pressure_Pa,
Sol25 = HESTIA_envirotox_cfs$Sol_mg.L,
Kdoc = rep(NA, nrow(HESTIA_envirotox_cfs)),
KpSS = rep(NA, nrow(HESTIA_envirotox_cfs)),
KpSd = rep(NA, nrow(HESTIA_envirotox_cfs)),
KpSl = rep(NA, nrow(HESTIA_envirotox_cfs)),
kdegA = HESTIA_envirotox_cfs$KdegA,
kdegW = HESTIA_envirotox_cfs$KdegW,
kdegSd = HESTIA_envirotox_cfs$KdegSd,
kdegSl = HESTIA_envirotox_cfs$KdegSl,
kdissP = rep(NA, nrow(HESTIA_envirotox_cfs)),
kdisswheat = rep(NA, nrow(HESTIA_envirotox_cfs)),
kdissrice = rep(NA, nrow(HESTIA_envirotox_cfs)),
kdisstomato = rep(NA, nrow(HESTIA_envirotox_cfs)),
kdissapple = rep(NA, nrow(HESTIA_envirotox_cfs)),
kdisslettuce = rep(NA, nrow(HESTIA_envirotox_cfs)),
kdisspotato = rep(NA, nrow(HESTIA_envirotox_cfs)),
avlogEC20 = HESTIA_envirotox_cfs$HC20,
ED50inhnoncanc = rep(NA, nrow(HESTIA_envirotox_cfs)),
ED50ingnoncanc = rep(NA, nrow(HESTIA_envirotox_cfs)),
ED50inhcanc = rep(NA, nrow(HESTIA_envirotox_cfs)),
ED50ingcanc = rep(NA, nrow(HESTIA_envirotox_cfs)),
BAFroot = rep(NA, nrow(HESTIA_envirotox_cfs)),
BAFleaf = rep(NA, nrow(HESTIA_envirotox_cfs)),
BTFmeat = rep(NA, nrow(HESTIA_envirotox_cfs)),
BTFmilk = rep(NA, nrow(HESTIA_envirotox_cfs)),
BAFfish = HESTIA_envirotox_cfs$BAF_L.Kg,
Flagged.Interim.Characterization.factor = rep(NA, nrow(HESTIA_envirotox_cfs)),
...41 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...42 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...43 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...44 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...45 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...46 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...47 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...48 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...49 = rep(NA, nrow(HESTIA_envirotox_cfs)),
USEtox.internal.identification.number = rep(NA, nrow(HESTIA_envirotox_cfs))
)

write.csv(USEtox_input, "../results/HESTIA_HC20_dataset_USEtox_format.csv", row.names = F)

```
