---
title: "Merging_Databases"
author: "Oskar Nyberg"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
  library(tidyverse)
  library(readxl)
```


## Reading source code from the "HESTIA_HC20_DB.Rmd"-document is required to make the following code functional.


## Merging HESTIA Tox DB with the EnviroTox DB.
The code below adds Envirotox "on top of" the HESTIA DB, Prioritizing HESTIA CASRN and Species, while adding CAS and species' unique to the Envirotox DB.

```{r Envirotox}
# Specifying which columns i want to keep in the main dataset
common_name_elements <- c(
  "CAS.Number", "DB", "version", "Database", "Species", "Genus", 
  "Taxonomy.Group", "Medium", "Effect", "Value.mg_l", "AcuteChronic", 
  "Endpoint", "Endpoint_conv", "Time.Hours",
  "EC10eq", "EC10eq_Chronic", "EC10eq_Acute", "No.Extrapolations", "Source")

# Reading in the Envirotox database (created in ../code/Envirotox_DB.Rmd)
EnviroTox_toxDB_EC10eq <- read.csv("../data/RAW_DB/EnviroTox_toxDB_EC10eq.csv")

# filtering out the unique Envirotox records
EnviroTox_Q_dat <- rbind(
  EnviroTox_toxDB_EC10eq %>% 
    # Removing all species taxonomic annotations given as "Genus sp."
    filter(!grepl("sp\\.", Species)) %>% 
    # selecting all CAS from `EnviroTox_substDB` that are NOT overlapping.
    filter(!CAS.Number %in% Q_dat$CAS.Number),
  EnviroTox_toxDB_EC10eq %>% 
    # Removing all species taxonomic annotations given as "Genus sp."
    filter(!grepl("sp\\.", Species)) %>% 
    # selecting all species that are unique to the `Q_dat` DB- 
    filter(CAS.Number %in% Q_dat$CAS.Number,
         !Species %in% Q_dat$Species))  %>% 
  select(common_name_elements)

# Adjusting the content of the HESTIA database
HESTIA_BASE_dat <- read.csv("../results/HESTIA_EC10eq_DB.csv") %>% 
  rename(Medium = Media.type) %>% 
  unite(Source, c("Title", "Author", "Year", "Reference.source", "URL", "DOI"), sep = ";", na.rm = TRUE) %>% 
  select(common_name_elements)

# This is the raw database with HESTIA as BASE and Envirotox as "filler"
HESTIA_BASE_EnviroTox_FILL <- rbind(HESTIA_BASE_dat, EnviroTox_Q_dat)

write.csv(HESTIA_BASE_EnviroTox_FILL, "../results/FINAL_HESTIA_BASE_EnviroTox_FILL.csv", row.names = F)
write.csv(HESTIA_BASE_EnviroTox_FILL, "../HESTIA_Shiny/data/FINAL_HESTIA_BASE_EnviroTox_FILL.csv", row.names = F)
```

## Merging all physchem & subst. use-type annotations 
```{r MergingPhyschem}
FULL_PHYSCHEM <- rbind(
  NEW_PHYSCHEM,
  read.csv("../data/excel_references/Envirotox_physchem.csv") %>% 
    # Duplicates in the two datasets are present. Filtering out these here.
    filter(!CAS.Number %in% NEW_PHYSCHEM$CAS.Number)) %>% 
  mutate(KdegA = case_when(KdegA == 0 ~ as.numeric(NA), TRUE ~ KdegA))
```


## Calculating HC50_EC50 for the HESTIA_envirotox Database for backwards compatibility with previous LCA-tox versions.
Dependency on `../code/HC50_calculation_function.R`-file with the defined function. 
This operation requires Physchem/subst.use-group data to be joined with the toxicological data.
```{r HC50EC50}
source("../code/HC50_calculation_function.R")

# Attaching the Physchem data to HESTIA dataset and outputting a df with HC50 data
HESTIA_HC50_calc <- HC50_calc_function(HESTIA_BASE_EnviroTox_FILL %>% left_join(x = ., y = FULL_PHYSCHEM, by = "CAS.Number"))
```

## HC20EC10eq calculations for HESTIA_BASE_EnviroTox_FILL

```{r HC20Calculations}
source("../code/HC20_calculation_function.R")

HESTIA_HC20_dataset <- HCx_calculator(HESTIA_BASE_EnviroTox_FILL, HCx = 20)

write.csv(HESTIA_HC20_dataset, "../results/HESTIA_HC20_dataset.csv", row.names = F)
```

## Running the weighted uncertainty model
```{r nlsModel, warning=FALSE, message=FALSE}
# Dataset to analyze is object `HESTIA_BASE_EnviroTox_FILL`

# load the nls function
 source("../code/least_squares_fit_model_code.R")
 # Running the operation with the HESTIA toxicity data set, generating a list object
system.time(nls_output <- nls_across_all(dataset = HESTIA_BASE_EnviroTox_FILL, Plot_output = "NO", MC_n = 1e+05, Plot_destination = "../figures/SSD_plots"))

#Transform into a dataframe - removing the list of lists containing full nls-output
nls_output_df <- as.data.frame(do.call(cbind, nls_output[1:16])) %>%
 mutate(across(c(2:15), ~ as.numeric(.x)),
        CAS.Number = as.factor(CAS.Number),
        status = as.character(status)) #%>% 
  # mutate(status = case_when(status == "warning: step factor reduced below 'minFactor'" ~ "Not enough data",
  #                           status == "singular gradient matrix at initial parameter estimates" ~ "One species record",
  #                           TRUE ~ status))

write.csv(nls_output_df, "../results/nls_output_df.csv", row.names = F)
write.csv(nls_output_df, "../HESTIA_Shiny/data/nls_output_df.csv", row.names = F)

nls_output_df <- read.csv("../results/nls_output_df.csv")

```


## Merging Physchem & Tox data
```{r FinalDataset}
HESTIA_envirotox_cfs <- left_join(
    x = FULL_PHYSCHEM, 
    y = HESTIA_HC20_dataset,
    by = "CAS.Number") %>% 
# Selection for the SHINY-app
  select(c(
    CAS.Number, PesticideAI_name, CanonicalSMILES, 
    MW.g.mol, pKaChemClass, pKa.gain, pKa.loss, Kow_L.L, Koc_L.kg_MCI,
    kH25C_Pa.m3.mol, Vapor.Pressure_Pa, Sol_mg.L,
    KdegA, KdegW, KdegSd, KdegSl, HC20, BAF_L.Kg, 
    mean_EC10eq, Sd_EC10eq, HC20EC10eq,
    Molecular.formula, Predefined.substance.type, Substance_type, Group, Subgroup, Heavy.Metals, Halogenated, definition, n_data_recs, n_species, sum_extrapolations, n_Taxonomy.Group 
    )
  ) %>% 
  mutate(across(contains("Kdeg"), ~ formatC(., format = "e")))

write.csv(HESTIA_envirotox_cfs, "../HESTIA_Shiny/data/HESTIA_envirotox_cfs.csv", row.names = F)

```

I'm reading in the USEtox Organics database names, to be able to export an identical list of data for easier USEtox input
```{r}

USEtox_input_names <- names(read_xlsx("../data/Impact assessment/USEtox_substance_data_organics.xlsx", sheet = "Substance data", skip = 1))
```

Creating an output file with the USEtox 2.1 database input structure 
Defining a data frame shaped like the USEtox-input database

```{r}
USEtox_input <- data.frame(
 ...1 = seq(1, nrow(HESTIA_envirotox_cfs), 1),
CASRN = HESTIA_envirotox_cfs$CAS.Number,
Name = HESTIA_envirotox_cfs$PesticideAI_name,
PesticideTargetClass = HESTIA_envirotox_cfs$Group,
PesticideChemClass = HESTIA_envirotox_cfs$Substance_type,
MW = as.numeric(HESTIA_envirotox_cfs$MW.g.mol),
pKaChemClass = HESTIA_envirotox_cfs$pKaChemClass,
pKa.gain = HESTIA_envirotox_cfs$pKa.gain,
pKa.loss = HESTIA_envirotox_cfs$pKa.loss,
KOW = HESTIA_envirotox_cfs$Kow_L.L,
Koc = HESTIA_envirotox_cfs$Koc_L.kg_MCI,
KH25C = HESTIA_envirotox_cfs$kH25C_Pa.m3.mol,
Pvap25 = HESTIA_envirotox_cfs$Vapor.Pressure_Pa,
Sol25 = HESTIA_envirotox_cfs$Sol_mg.L,
Kdoc = rep(NA, nrow(HESTIA_envirotox_cfs)),
KpSS = rep(NA, nrow(HESTIA_envirotox_cfs)),
KpSd = rep(NA, nrow(HESTIA_envirotox_cfs)),
KpSl = rep(NA, nrow(HESTIA_envirotox_cfs)),
kdegA = HESTIA_envirotox_cfs$KdegA,
kdegW = HESTIA_envirotox_cfs$KdegW,
kdegSd = HESTIA_envirotox_cfs$KdegSd,
kdegSl = HESTIA_envirotox_cfs$KdegSl,
kdissP = rep(NA, nrow(HESTIA_envirotox_cfs)),
kdisswheat = rep(NA, nrow(HESTIA_envirotox_cfs)),
kdissrice = rep(NA, nrow(HESTIA_envirotox_cfs)),
kdisstomato = rep(NA, nrow(HESTIA_envirotox_cfs)),
kdissapple = rep(NA, nrow(HESTIA_envirotox_cfs)),
kdisslettuce = rep(NA, nrow(HESTIA_envirotox_cfs)),
kdisspotato = rep(NA, nrow(HESTIA_envirotox_cfs)),
avlogEC20 = HESTIA_envirotox_cfs$HC20,
ED50inhnoncanc = rep(NA, nrow(HESTIA_envirotox_cfs)),
ED50ingnoncanc = rep(NA, nrow(HESTIA_envirotox_cfs)),
ED50inhcanc = rep(NA, nrow(HESTIA_envirotox_cfs)),
ED50ingcanc = rep(NA, nrow(HESTIA_envirotox_cfs)),
BAFroot = rep(NA, nrow(HESTIA_envirotox_cfs)),
BAFleaf = rep(NA, nrow(HESTIA_envirotox_cfs)),
BTFmeat = rep(NA, nrow(HESTIA_envirotox_cfs)),
BTFmilk = rep(NA, nrow(HESTIA_envirotox_cfs)),
BAFfish = HESTIA_envirotox_cfs$BAF_L.Kg,
Flagged.Interim.Characterization.factor = rep(NA, nrow(HESTIA_envirotox_cfs)),
...41 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...42 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...43 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...44 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...45 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...46 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...47 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...48 = rep(NA, nrow(HESTIA_envirotox_cfs)),
...49 = rep(NA, nrow(HESTIA_envirotox_cfs)),
USEtox.internal.identification.number = rep(NA, nrow(HESTIA_envirotox_cfs))
)

write.csv(USEtox_input, "../results/HESTIA_HC20_dataset_USEtox_format.csv", row.names = F)

```

## Introduction
```{r}
# EnviroTox_DB <- readxl::read_xlsx(
#   path = "data/Envirotox_2023-01-10/envirotox_DB.xlsx", sheet = "test", col_names = T, col_types = c(
#     "skip", rep("text", times = 4), "guess", rep("text", times = 4), "guess", "guess", "logical", rep("text", times = 3), "numeric") ) %>% 
#   rename(
#    Endpoint = `Test statistic`,
#    AcuteChronic= `Test type`,
#    Species = `Latin name`,
#    Value.mg_l = `Effect value`,
#    Time.Hours = `Duration (hours)`) %>% 
#   mutate(
#     DB = "EnviroTox",
#     CAS.Number = as.cas(`original CAS`),
#     AcuteChronic = case_when(
#       AcuteChronic == "A" ~ "Acute",
#       TRUE ~ "Chronic"))
```

Additional data are sourced from EnviroTox, a curated ecotoxicological database for `r nrow(EnviroTox_DB %>% distinct(CAS.Number))` substances [@connors_2019].

## Methods

### Including additional data sources
We merge the curated ecotoxicological effect dataset with selected (non-overlapping) records from the EnviroTox toxicological database. The EnviroTox toxicological database by Connors et al., [-@connors_2019] is a curated aquatic toxicological database containing a large set of toxicological data from a broad range of potentially toxic substances. Prior to merging the two databases, curation of the EnviroTox database was performed accordingly:  
Toxicological endpoints were selected and grouped as for HESTIA database (Table STX), Acute and Chronic test duration definitions were performed according to taxonomic group and duration of experiment. The EnviroTox team have thoroughly refined acute/chronic exposure selection methodology yet for the current purpose we chose to categorize acute/chronic definitions in a harmonized way according to the methodology described above. Taxonomic information from the EnviroTox database was slightly revised to match the HESTIA dataset (e.g minor spelling corrections). With harmonized acute/chronic definitions, identical taxonomic descriptions and endpoint conversions, extrapolation factors for EC10eq conversions were added to effect data as for the original dataset. Lastly, records with duplicates across all rows were removed. Data sources overlap across the HESTIA and EnviroTox datasets, and records are selected from EnviroTox when 1) all records with CASRN unique to EnviroTox, and 2) records where species, but not CASRN, are unique to EnviroTox, for details, see SI 4. 

```{r}
# Merging all physchem & subst. use-type annotations 
FULL_PHYSCHEM <- rbind(
  NEW_PHYSCHEM,
  read.csv("data/excel_references/Envirotox_physchem.csv") %>% 
    # managed to get some duplicates in here. Could fix by redoing physchem query at OECD QSAR Toolbox in Envirotox wrangle, but also just down-prioritize (filter out) those duplicates here.
    filter(!CAS.Number %in% NEW_PHYSCHEM$CAS.Number))
```

## Results
As a result of selecting records unique to the Envirotox database and curating the data as previously described, `r nrow(HESTIA_BASE_EnviroTox_FILL %>% filter(DB == "EnviroTox"))` records are added to the final dataset, summarized in Table \@ref(tab:Envirotox).  

```{r Envirotox}
# Make sure to be able to reference to this output!!
Envirotox_summary_table <- flextable(HESTIA_BASE_EnviroTox_FILL %>%
  filter(DB == "EnviroTox") %>% 
  group_by(Taxonomy.Group, Endpoint_conv, AcuteChronic) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = c(Endpoint_conv, AcuteChronic), values_from = n) %>% 
  mutate(Sum_Species = sum(across(contains("_")), na.rm = TRUE)) %>%
  ungroup() %>%
  bind_rows(summarise(
    .,
    across(where(is.numeric), ~sum(., na.rm = T)),
    across(where(is.character), ~as.character("Total") ))))

Envirotox_summary_table %>% 
  set_caption("Number of records from EnviroTox dataset joined to HESTIA by taxonomic group and endpoint.") %>% 
  fontsize(., size = 10, part = "all") %>% 
  flextable::set_table_properties(layout = "autofit") %>% 
  flextable::set_header_labels(
    Taxonomy.Group = "Taxonomy group", 
    EC10_Chronic = "EC10 chronic",  
    EC50_Chronic = "EC50 chronic",  
    NOEC_Chronic = "NOEC chronic",  
    EC10_Acute = "EC10 acute",  
    EC50_Acute = "EC50 acute",  
    NOEC_Acute = "NOEC acute",  
    Sum_Species = "Total per taxa") 

```

