---
title: "Physicochemical_properties_Envirotox"
author: "Oskar Nyberg"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(verbose = TRUE, warning = FALSE)
    library(rmarkdown)
    library(xlsx)
    library(webchem)
    library(tidyverse)
```

Select all the Envirotox data lacking physchem or pesticide info.
```{r}
# Importing a vector of CAS.Numbers where Physicochemical and Pesticide use data is available 
 # main_CAS <- read.csv("../data/QSAR_Toolbox_physchem_data_2023-03-24_RAW.csv", header = T, sep = "\t", na.strings = "", fileEncoding = "UTF-16LE",  stringsAsFactors = FALSE) %>% 
 #  pull(CAS.Number)

# # no need to read this dataset, since it's loaded at the bottom along with PubChemId!
# # Select CAS and SMILES unique to the Envirotox database
#   Envirotox_no_physchem <- readxl::read_xlsx(path = "../data/Envirotox_2023-01-10/envirotox_DB.xlsx", sheet = "substance", col_names = T, col_types = c("skip", "guess", rep("text", times = 2), rep("skip", times = 19))) %>%
#     mutate(CAS.Number = as.cas(`original CAS`)) %>%
#     rename(CanonicalSMILES = `Canonical SMILES`) %>%
#     select(CAS.Number, CanonicalSMILES, `Chemical name`) %>%
#     separate(`Chemical name`, into = c("PesticideAI_name", "Envirotox_name"), sep = ";", remove = T, convert = TRUE) #%>%
#     #filter(!CAS.Number %in% main_CAS)
# 
# # So the SMILES configurations seem a bit wierd. Perhaps it is a ChemSpider format with '@' etc.
# Envirotox_no_physchem$cid <- get_cid(Envirotox_no_physchem$CanonicalSMILES, from = "smiles", domain = "compound", match = "all", verbose = T)$cid
# Envirotox_no_physchem <- Envirotox_no_physchem %>%
#   mutate(cid = case_when(CAS.Number == "30171-80-3" ~ "3035339",
#                          CAS.Number == "99685-96-8" ~ "123591",
#                          CAS.Number == "26967-76-0" ~ "75628",
#                          CAS.Number == "9000-30-0" ~ "441155876",
#                          TRUE ~ cid
#                      ))
#write.csv(Envirotox_no_physchem, "../data/excel_references/Envirotox_no_physchem.csv", row.names = F)

# Writing an output that will generate Physchem data for the EnviroTox dataset, the records that are missing.
# write.table(Envirotox_no_physchem[, c("CAS.Number", "CanonicalSMILES")], "../data/excel_references/Envirotox_no_physchem_2.txt", col.names = F, row.names = F, na = "", sep = "\t", quote = FALSE)

Envirotox_no_physchem_CID <- read.csv("../data/excel_references/Envirotox_no_physchem.csv") %>% 
  rename(PubChemId = cid) %>% 
  mutate(PubChemId = as.character(PubChemId))
```


## USEPA Comptox categorized lists
```{r USEPA-Comptox}
# Make a vector of file names that are to be read in upcoming function
USEPA_filenames <- list.files(path = "../data/USEPA - Comptox_categories", pattern = "Chemical List", full.names = TRUE)

# Function loading all the USEPA_COMPTOX substance information annotation
# need to load 16 files with a lot of specific stuff to read in. 
USEPA_CHEMLIST <- do.call(rbind, lapply(USEPA_filenames, function(x){
  read.csv(x, header = T, sep = ",", colClasses = c("NULL", rep("character", 7), rep("numeric", 7), "character", rep("numeric", 2))) %>%
    select(CASRN, PREFERRED.NAME, IUPAC.NAME, SMILES,  MOLECULAR.FORMULA) %>%
    mutate(Source = "USEPA_COMPTOX",
           Sub_source = paste(as.character(            # Inserting part of substance information in column "Group", based on file name
             gsub("../data/USEPA - Comptox_categories/Chemical List ", "",
              gsub("-2023-01-20.csv", "",
                gsub("WIKI", "Wikipedia", 
                  gsub(",.*", "", x)))))),
          Group = paste(as.character(                  # Inserting part of substance information in column "Group", based on file name
             gsub("../data/USEPA - Comptox_categories/Chemical List ", "",
              gsub("-2023-01-20.csv", "",
                gsub("WIKI", "Wikipedia", 
                  sub(".*?,", "", x)))))))
            }
    )
  ) %>% 
  pivot_wider(id_cols = c(CASRN, PREFERRED.NAME, IUPAC.NAME, SMILES,  MOLECULAR.FORMULA), names_from = Sub_source, values_from = Group) %>% 
  unnest(cols = everything()) %>% 
  filter(!grepl("NOCAS", CASRN)) 

write.csv(USEPA_CHEMLIST, "../data/excel_references/USEPA_CHEMLIST_2023-01-20.csv", row.names = FALSE)

USEPA_CHEMLIST <- USEPA_CHEMLIST %>% 
  rename(CAS.Number = CASRN) %>% 
  select(-c("PREFERRED.NAME", "IUPAC.NAME", "SMILES", "MOLECULAR.FORMULA")) %>% 
  group_by(CAS.Number, DRUGBANK, EPAPCS, HEALTHY_BUILDING_NETWORK, NORMAN_ITN, OPPIN, PPDB) %>% 
  summarize(
      USEPA = paste(USEPA, collapse = "; "),
      Wikipedia = paste(Wikipedia, collapse="; ")) %>% 
  ungroup() %>% 
  mutate(
    across(USEPA:Wikipedia, ~ na_if(., "NA")),
    across(USEPA:Wikipedia, ~ na_if(., "NA; NA")))

#write.xlsx(USEPA_CHEMLIST, "data/excel_references/USEPA_CHEMLIST_2023-01_20.xlsx", sheetName = "Data1", col.names = T, append = T, showNA = F)

```

**US EPA ECOTOX DB** (locally downloaded file) - accessed 2022-march-10 - is used to extract information on "ecotox_group" for substances. This file has `r nrow(USEPA_Ecotox_Chem_DB)` rows and is eerily similar to HESTIA's PesticideAI.csv file which this whole work is based off of. 

```{r USEPA ECOTOX DB}
# US EPA ECOTOX Chemical Type Annotations
USEPA_Ecotox_Chem_DB <- read.delim("../../../Backup HESTIA/USEPA-ECOTOX DB/ecotox_ascii_03_10_2022/validation/chemicals.txt", header = T, sep = "|") %>% 
  mutate(CAS.Number = as.cas(cas_number))%>% 
  select(-cas_number)%>% 
  mutate(
    dtxsid = gsub("None", as.character(NA), dtxsid),
    ecotox_group = gsub("None", as.character(NA), ecotox_group),
    ecotox_group = sub("PPCPs", "PPCP", ecotox_group),
    ecotox_group = sub("EDCs", "EDC", ecotox_group),
    ecotox_group = sub("PAHs", "PAH", ecotox_group),
    across(starts_with("BCPC"), ~ case_when(.x == "NA" ~ as.character(NA), TRUE ~ .x))) %>% 
  rename(USEPA_ecotox_group = ecotox_group)


```

Additional Database queries comes from the following repositories: British Compendium of Pesticide Common Names (BCPC) which is accessed via the `webchem` package []. 
```{r BCPC}
# Query the BCPC Compendium of Pesticide Common Names https://pesticidecompendium.bcpc.org
  # Query and code has been made as: 

#Pest_query <- bcpc_query(Envirotox_no_physchem_CID$CAS.Number, from = "cas", verbose = T)
#saveRDS(Pest_query, "../data/excel_references/Pest_query_enviro.txt")
Pest_query <- readRDS("../data/excel_references/Pest_query_enviro.txt")
#Pest_drop.na <- Pest_query[!is.na(Pest_query)] # Removing NAs list objects. not really needed here.

Pest_rbind <- as.data.frame(do.call(rbind, Pest_query)) # Flatten the nested lists into one comprehensive data frame
Pest_rbind$Query_CAS <- rownames(Pest_rbind)

rownames(Pest_rbind) = NULL # Removing rownames

Pest_rbind[1:12] <- lapply(Pest_rbind[1:12], as.character) # Coercing all data into character. Previously each cell was defined as a list.

Pest_info <- Pest_rbind %>% # selecting the cols i want to add in the final output
  mutate(
  activity = gsub("cides([[:alpha:]])", "cides \\1", activity),
  activity = gsub("repellents([[:alpha:]])", "repellents \\1", activity),
  activity = gsub("bird repellents", "bird_repellents", activity),
  activity = gsub("mammal repellents", "mammal_repellents", activity),
  activity = gsub("plant growth regulators", "plant_growth_regulators", activity),
  activity = gsub("\\(.*?\\)", "", activity)
       ) %>% 
  rename(
    CAS.Number = Query_CAS,
    BCPC_formula = formula,
    BCPC_cname = cname,
    BCPC_iupac_name = iupac_name,
    BCPC_activity = activity,
    BCPC_subactivity = subactivity,
    BCPC_source_url = source_url) %>% 
  mutate(BCPC_pesticide = case_when(grepl("icides", BCPC_activity) ~ "Pesticide", TRUE ~ as.character(NA))) %>% 
  separate(BCPC_activity, into = c("BCPC_activity_1", "BCPC_activity_2", "BCPC_activity_3", "BCPC_activity_4", "BCPC_activity_5"), sep = " ", remove = T) %>%
  unite("BCPC_activity", c("BCPC_activity_1", "BCPC_activity_2", "BCPC_activity_3", "BCPC_activity_4", "BCPC_activity_5"), sep = "; ", na.rm = T, remove = T) %>% 
  mutate_all( ~ na_if(., '')) %>% 
  mutate_all( ~ na_if(., 'NA')) %>% 
  select(-c(inchikey, inchi))
```

Looking up substances' classification from from the ATC_class list as well as the **ChEMBL** (https://www.ebi.ac.uk/chembl/) - a manually curated database of bio-active molecules with drug-like properties. It brings together chemical, bioactivity and genomic data to aid the translation of genomic information into effective new drugs.
```{r ChEMBL}
# Checking out the new ChEMBL service in latest Webchem update (as of early 2023 ;)
  # ATC_classes <- chembl_atc_classes(verbose = T)
  # write_rds(ATC_classes, "data/excel_references/ATC_classes.txt")
ATC_classes <- read_rds("../data/excel_references/ATC_classes.txt")

ChEMBL_agents <- ATC_classes %>% 
    filter(level1 %in% c("P", "J"))

# defining names to be used in the CIR query
names <- ChEMBL_agents %>% 
    pull(who_name)

# ChEMBL_CID <- get_cid(names, from = "name", domain = "compound", match = "all", verbose = TRUE)
  # write_rds(ChEMBL_CID, "../data/excel_references/ChEMBL_CID.txt")
ChEMBL_CID <- read_rds("../data/excel_references/ChEMBL_CID.txt") %>% 
  rename(who_name = query,
         PubChemId = cid) %>%
  filter(!is.na(PubChemId))

# Merging with the ATC_CLASS df
ChEMBL_data <- ChEMBL_CID %>% 
    left_join(x = .,
              y = ChEMBL_agents,
              by = "who_name") %>% 
    rename_at(vars(matches("level")), ~ paste0("ATC_", .))%>% 
    rename(WHO_ATC_name = who_name) %>% 
    # some substances are classified as multiple properties. Dealing with this here
    mutate(
    # Also, selecting relevant columns and renaming these for better descriptions below
      ATC_Type = case_when(grepl(paste(c("ANTIBACTERIALS", "ANTIMYCOBACTERIALS"), collapse = "|"), ATC_level2_description) ~ str_to_title("ANTIBACTERIAL"),
      grepl("ANTIVIRALS", ATC_level2_description) ~ str_to_title("ANTIVIRAL"),
      grepl("ANTIPARASITIC PRODUCTS, INSECTICIDES AND REPELLENTS", ATC_level1_description) ~ str_to_title("ANTIPARASITIC; INSECTICIDES; REPELLENTS"))) %>% 
    rename(ATC_Description = ATC_level4_description) %>% 
    select(WHO_ATC_name, PubChemId, ATC_Type, ATC_Description) %>% 
    distinct() %>% 
    # Solution to the duplicate columns. They have been merged together!
    group_by(PubChemId, WHO_ATC_name) %>% summarize(
      ATC_Type=paste(ATC_Type, collapse=","),
      ATC_Description=paste(ATC_Description, collapse=","))
```

## The ChEBI database has some cool information in the "parents" subsection. 
https://www.ebi.ac.uk/chebi/init.do 
I scrape the database for all data available current on substance list. 
```{r ChEBI database}
# dependency:`CAS_CID_list` object loaded in the first chunk

#HESTIA_ChEBI_name <- get_chebiid(CAS_CID_list$CAS.Number, from = "registry numbers", match = "best", stars = "all", verbose = T) # OK, Done
#write_rds(HESTIA_ChEBI_name, "HESTIA_Chem_Chebi_IDs.txt")
HESTIA_ChEBI_name <- read_rds("../data/excel_references/HESTIA_Chem_Chebi_IDs.txt")

#HESTIA_ChEBI_all_list <- chebi_comp_entity(HESTIA_ChEBI_name$chebiid, verbose = T) #OK, Done
#write_rds(HESTIA_ChEBI_all_list, "HESTIA_Chem_all_ChEBI_data.txt")
HESTIA_ChEBI_all_list <- read_rds("../data/excel_references/HESTIA_Chem_all_ChEBI_data.txt")

# Starting with the wrangling of lists:
HESTIA_ChEBI_all_list_drop.na <- HESTIA_ChEBI_all_list[!is.na(HESTIA_ChEBI_all_list)] # Removing NAs list objects
#HESTIA_ChEBI_all_list__flat <- as.data.frame(do.call(rbind, HESTIA_ChEBI_all_list_drop.na))
HESTIA_ChEBI_all_list_parents <- do.call("rbind", lapply(HESTIA_ChEBI_all_list_drop.na, "[[", 10) ) # Flattening the "Parents" data frame with substance information
HESTIA_ChEBI_all_list_parents$ChEBId <- rownames(HESTIA_ChEBI_all_list_parents) # Creating a substance name column
rownames(HESTIA_ChEBI_all_list_parents) = NULL # Removing rownames

# list of pesticides;
HESTIA_ChEBI_definitions <- HESTIA_ChEBI_all_list_parents %>% 
    #filter(type == "has role") %>% 
    separate(col = ChEBId, into = c("CheBId", "row.nr"), sep = "\\.", remove = T) %>% 
    mutate(type = case_when(chebiId %in% c("CHEBI:38805", "CHEBI:22153", "CHEBI:39215", "CHEBI:22583", "CHEBI:51076", "CHEBI:86328", "CHEBI:33289",
                                           "CHEBI:23092", "CHEBI:39276", "CHEBI:24527", "CHEBI:24852", "CHEBI:33904", "CHEBI:25491", "CHEBI:25943",
                                           "CHEBI:167183", "CHEBI:136643", "CHEBI:33288", "CHEBI:25944", "CHEBI:38601", "CHEBI:38656", "CHEBI:9442",
                                           "CHEBI:73333", "CHEBI:38956"
                                           ) ~ "pesticide",
                            
                            chebiId %in%  c("CHEBI:33286", "CHEBI:71692"
                                            ) ~ "agrochemical",
                            
                            chebiId %in%  c("CHEBI:33281", "CHEBI:9337", "CHEBI:87230", "CHEBI:87228", "CHEBI:86324", "CHEBI:86322", "CHEBI:26179",  
                                            "CHEBI:26177", "CHEBI:25807", "CHEBI:25105", "CHEBI:22507", "CHEBI:156449", "CHEBI:15369", "CHEBI:23066",
                                            "CHEBI:23765", "CHEBI:35358", "CHEBI:23007", "CHEBI:25558", "CHEBI:25605", "CHEBI:25903", "CHEBI:27933",
                                            "CHEBI:39217", "CHEBI:49318", "CHEBI:49322", "CHEBI:86478", "CHEBI:87211", "CHEBI:88225", "CHEBI:36043"
                                            ) ~ "antimicrobial",
                            
                            chebiId %in%  c("CHEBI:33282", "CHEBI:36047") ~ "antibacterial",
                            chebiId %in%  c("CHEBI:22587", "CHEBI:36044") ~ "antiviral",
                            chebiId %in%  c("CHEBI:39094", "CHEBI:35444"
                                            ) ~ "nematocide",
                            
                            chebiId %in%  c("CHEBI:33287", "CHEBI:16199"
                                            ) ~ "fertilizer",
                            
                            chebiId %in%  c("CHEBI:82599", "CHEBI:25540", "CHEBI:25705", "CHEBI:26409", "CHEBI:38455", "CHEBI:38611", "CHEBI:38804",
                                            "CHEBI:39116", "CHEBI:39190", "CHEBI:39213", "CHEBI:136644", "CHEBI:22917", "CHEBI:23457", "CHEBI:25708",
                                            "CHEBI:25715", "CHEBI:38060", "CHEBI:38461", "CHEBI:38488", "CHEBI:38494", "CHEBI:38577", "CHEBI:39090",
                                            "CHEBI:39167", "CHEBI:39191", "CHEBI:39208", "CHEBI:39210", "CHEBI:39277", "CHEBI:39295", "CHEBI:39351",
                                            "CHEBI:39398", "CHEBI:39415"
                                            ) ~ "insecticide",
                            
                            chebiId %in%  c("CHEBI:60575", "CHEBI:60575", "CHEBI:133673", "CHEBI:15930", "CHEBI:137504"
                                            ) ~ "herbicide",
                            
                            chebiId %in%  c("CHEBI:36053", "CHEBI:38489", "CHEBI:38602", "CHEBI:38612", "CHEBI:38657", "CHEBI:38806", "CHEBI:38820",
                                            "CHEBI:39219", "CHEBI:39259", "CHEBI:39292", "CHEBI:39296", "CHEBI:39298", "CHEBI:39301", "CHEBI:39318",
                                            "CHEBI:39363", "CHEBI:39369", "CHEBI:39412"
                                            ) ~ "acaricide",
                            
                            chebiId %in%  c("CHEBI:24127", "CHEBI:87208", "CHEBI:87198", "CHEBI:87197", "CHEBI:87195", "CHEBI:87114", "CHEBI:87100",
                                            "CHEBI:87068", "CHEBI:87067", "CHEBI:87066", "CHEBI:87064", "CHEBI:87061", "CHEBI:87039", "CHEBI:87038",
                                            "CHEBI:87071", "CHEBI:87069", "CHEBI:87035", "CHEBI:87034", "CHEBI:87019", "CHEBI:87018", "CHEBI:87015",
                                            "CHEBI:60600", "CHEBI:87113", "CHEBI:86327", "CHEBI:86494", "CHEBI:86488", "CHEBI:86487", "CHEBI:86484",
                                            "CHEBI:86482", "CHEBI:35718", "CHEBI:38819", "CHEBI:86417", "CHEBI:86440", "CHEBI:86478", "CHEBI:86485",
                                            "CHEBI:87013", "CHEBI:87021", "CHEBI:87022", "CHEBI:87025", "CHEBI:87026", "CHEBI:87027", "CHEBI:87029",
                                            "CHEBI:87036", "CHEBI:87037", "CHEBI:87101", "CHEBI:87110", "CHEBI:87127", "CHEBI:87128", "CHEBI:87134",
                                            "CHEBI:87135", "CHEBI:87207", "CHEBI:136645" 
                                            ) ~ "fungicide",
                            
                           TRUE ~ as.character(NA))) 

ChEBI_annot <-  HESTIA_ChEBI_definitions %>% 
                distinct(CheBId, .keep_all = T) %>% 
                select(chebiName, CheBId) %>%
                left_join(x =.,
                          y = HESTIA_ChEBI_definitions %>% 
                                filter(!grepl(paste(c("metabolite", "inhibitor"), collapse = "|"), chebiName),
                                       !is.na(type)) %>% 
                                select(-c(row.nr, status, cyclicRelationship, chebiName)) %>% 
                                pivot_wider(values_from = type, names_from = chebiId) %>% 
                                unnest(cols = everything() ) %>% 
                                unite("ChEBI_DB", 2:160, sep = ";", na.rm = T),
                          by = "CheBId") %>% 
                rename(chebiid = CheBId) %>% 
# a separate-and-clean exercise of the CheBI_annot object
    separate(ChEBI_DB, into = c("ChEBI_DB", "ChEBI_DB.1", "ChEBI_DB.2", "ChEBI_DB.3", "ChEBI_DB.4", "ChEBI_DB.5", "ChEBI_DB.6"), sep = ";", remove = TRUE) %>% 
    mutate(ChEBI_DB.1 = case_when(ChEBI_DB.1 == ChEBI_DB ~ as.character(NA),
              TRUE ~ ChEBI_DB.1),
           ChEBI_DB.2 = case_when(ChEBI_DB.2 == ChEBI_DB | 
                                  ChEBI_DB.2 == ChEBI_DB.1 ~ as.character(NA),
              TRUE ~ ChEBI_DB.2),
           ChEBI_DB.3 = case_when(ChEBI_DB.3 == ChEBI_DB | 
                                ChEBI_DB.3 == ChEBI_DB.1 |
                                ChEBI_DB.3 == ChEBI_DB.2 ~ as.character(NA),
              TRUE ~ ChEBI_DB.3),
           ChEBI_DB.4 = case_when(ChEBI_DB.4 == ChEBI_DB | 
                                ChEBI_DB.4 == ChEBI_DB.1 |
                                ChEBI_DB.4 == ChEBI_DB.2 |
                                ChEBI_DB.4 == ChEBI_DB.3 ~ as.character(NA),
              TRUE ~ ChEBI_DB.4),
           ChEBI_DB.5 = case_when(ChEBI_DB.5 == ChEBI_DB | 
                                ChEBI_DB.5 == ChEBI_DB.1 |
                                ChEBI_DB.5 == ChEBI_DB.2 |
                                ChEBI_DB.5 == ChEBI_DB.3 |
                                ChEBI_DB.5 == ChEBI_DB.4 ~ as.character(NA),
              TRUE ~ ChEBI_DB.5),
           ChEBI_DB.6 = case_when(ChEBI_DB.6 == ChEBI_DB | 
                                ChEBI_DB.6 == ChEBI_DB.1 |
                                ChEBI_DB.6 == ChEBI_DB.2 |
                                ChEBI_DB.6 == ChEBI_DB.3 |
                                ChEBI_DB.6 == ChEBI_DB.4 |
                                ChEBI_DB.6 == ChEBI_DB.5 ~ as.character(NA),
              TRUE ~ ChEBI_DB.6)
           ) %>% 
  unite("ChEBI_DB", 3:9, sep = ";", na.rm = T)

ChEBI_annot$ChEBI_DB <- sapply(strsplit(as.character(ChEBI_annot$ChEBI_DB), ';\\s*'), 
                     function(x) toString(sort(x, decreasing = T)))
      
# Flattening the "regnumbers" data frame with Registry numbers
# But it's "too" flat! I have lost the ChEBI ID nr.
HESTIA_ChEBI_all_list_regnumbers <- do.call("bind_rows", lapply(HESTIA_ChEBI_all_list_drop.na, "[[", 7) ) %>%  # The dplyr "bind_rows" function inserts NAs to force together dfs with uneven amount of columns!! 
                                      filter(type == "CAS Registry Number") %>% 
                                      distinct(data, .keep_all = T)

reg_nr <- as.data.frame(do.call(rbind, HESTIA_ChEBI_all_list_drop.na)) %>% 
    select(regnumbers)

reg_nr <- do.call("rbind", lapply(reg_nr$regnumbers, "[[", 1) )
reg_nr <- as.data.frame(reg_nr)
ChEBI_ID_CAS <- reg_nr %>% 
        mutate(across(everything(), ~ case_when(is.cas(.x) ~ .x, 
                                                   TRUE ~ as.character(NA)))) %>% 
        rename(CAS.Number = V1) %>% 
        mutate(CAS.Number = case_when(is.na(CAS.Number) ~ V2, 
                                      TRUE ~ CAS.Number),
               CAS.Number = case_when(is.na(CAS.Number) ~ V3, 
                                      TRUE ~ CAS.Number),
               CAS.Number = case_when(is.na(CAS.Number) ~ V4, 
                                      TRUE ~ CAS.Number),
               CAS.Number = case_when(is.na(CAS.Number) ~ V5, 
                                      TRUE ~ CAS.Number),
               CAS.Number = case_when(is.na(CAS.Number) ~ V6, 
                                      TRUE ~ CAS.Number),
               CAS.Number = case_when(is.na(CAS.Number) ~ V7, 
                                      TRUE ~ CAS.Number)
               )
  
ChEBI_ID_CAS$ChebId <- rownames(ChEBI_ID_CAS) 
rownames(ChEBI_ID_CAS) = NULL # Removing rownames

# Adding lot's of inchl and stuff... is it important?? NO. Not now..
p <- as.data.frame(do.call(rbind, HESTIA_ChEBI_all_list_drop.na)) %>% 
    select(properties)
p$ChEBId <- rownames(p)
rownames(p) = NULL # Removing rownames

pq <- do.call(rbind, p$properties) %>% 
      select(chebiid, chebiasciiname, definition, smiles, inchi, inchikey, mass)
rownames(pq) = NULL # Removing rownames

#####
# MERGING ALL THE ChEBI DATA#
#####
ChEBI_annot_merge <- ChEBI_annot %>% 
   mutate(
    ChEBI_DB = na_if(ChEBI_DB, ''),
    ChEBI_DB = gsub("\\,", ";", ChEBI_DB),
         ) %>% 
  left_join(x = ., 
            y = ChEBI_ID_CAS %>% 
                select(ChebId, CAS.Number) %>%
                mutate(ChebId = gsub("\\.", ":", ChebId)) %>% 
                rename(chebiid = ChebId), 
            by = "chebiid") %>% 
  left_join(x = ., 
            y = pq, 
            by = "chebiid") %>% 
  distinct(CAS.Number, .keep_all = TRUE)
```


```{r}
Envirotox_Comp_info <- Envirotox_no_physchem_CID %>% 
    # Merging US EPA ECOTOX Chemical Type Annotations with HESTIA CAS numbers list
    left_join(
      x = . ,
      y = USEPA_Ecotox_Chem_DB,
      by = "CAS.Number") %>% 
    # Merging BCPC Pest_info to the HESTIA_Comp_info
    left_join(
      x = ., 
      y = Pest_info, 
      by = "CAS.Number") %>%
    left_join(
      x = ., 
      y = ChEBI_annot_merge, 
      by = "CAS.Number") %>% 
    left_join(
      x = .,
      y = ChEMBL_data,
      by = "PubChemId") %>% 
    left_join(
      x = ., 
      y = USEPA_CHEMLIST,
      by = "CAS.Number") %>%
    select(c(
"CAS.Number", "CanonicalSMILES", "smiles", "dtxsid", # Chem_id's
"PesticideAI_name", "chebiasciiname", "chemical_name", "BCPC_cname", "WHO_ATC_name", "BCPC_iupac_name", # "chebiName",  # Names
"mass", "BCPC_formula", # Chem_props
"ATC_Type", "ATC_Description", "USEPA_ecotox_group", "BCPC_pesticide", "BCPC_activity", "ChEBI_DB", "DRUGBANK", "EPAPCS", 
"HEALTHY_BUILDING_NETWORK", "NORMAN_ITN", "OPPIN", "PPDB", "USEPA", "Wikipedia", # Pest IINFO
"PubChemId", "chebiid",  "definition", "inchi", "inchikey", # ID_codes
"BCPC_source_url", 
    ))
```

## Substance Type Annotations
Creating an annotation on substances being 1. Inorganic/Organic, 2. containing Heavy metals, and 3. being Halogenated.
```{r}
# Halogenated: contains c(F, Cl, Br, I)

# Investigating smiles that contain C in the smiles (Cu, Cr, Co),yet excluding such elements with a density <5g/cm3 and Carbon (C). !
# Creatng a vector of substances that I can filter against.
# This vector is based on 3 operations: 
  # 1. Looking up all SMILES containing elements with the letter "C" and filtering out all possible carbon components
  # 2. Looking up all elements not containing carbon whatsoever
  # 3. Merging both lists as CAS.numbers to filter out.
# 332 substances are identified. Although substances without SMILES configuration will be remain in the data set!

Inorganic_vector_x <- rbind(Envirotox_Comp_info %>% 
                              distinct(CAS.Number, .keep_all = T) %>% # this filter detects all SMILES 
                              filter(str_detect(CanonicalSMILES, paste(c("Cd", "Cs", "Co", "Cs", "Cr", "Cu", "Cl", "Ca"), collapse = "|")), # Making sure to select all substances with elements containing a "C" or "c" in SMILES
                                     !str_detect(CanonicalSMILES, paste(c("CC", "c","C+[0-9]", "C+\\(", "C+\\)", "C+\\[", "C+\\]", "C+\\#", "C+\\="), collapse="|"))),
                          Envirotox_Comp_info %>% # This filter detects all SMILES without Carbon.
                              distinct(CAS.Number, .keep_all = T) %>% 
                              filter(!str_detect(CanonicalSMILES, paste(c("c","C"), collapse = "|")))) %>% 
                          pull(CAS.Number)

HM_vector_x <- Envirotox_Comp_info %>% # Just as in the operation above with an inorganic annotation, This creates a vector of CAS.Numbers to annotate Heavy metals presence in the smiles configuration
            distinct(CAS.Number, .keep_all = T) %>% 
            filter(str_detect(CanonicalSMILES, paste(read.xlsx("../data/excel_references/Elements_list.xlsx", sheetName = "Sheet1", header = T) %>% # Using a list of elements (From Wikipedia) with annotations of Elements and their density
                                                      mutate(Density = as.numeric(Density)) %>% 
                                                      filter(Density >=5 & Element >20) %>% 
                                                      pull(Abbr), 
                                                      collapse = "|"))) %>% 
            pull(CAS.Number)

Halogenated_vector_x <- Envirotox_Comp_info %>% # Just as in the operation above with an inorganic annotation, This creates a vector of CAS.Numbers to annotate Halogens presence in the SMILES configuration
                        distinct(CAS.Number, .keep_all = T) %>% 
                        filter(str_detect(CanonicalSMILES, paste(c("F", "Cl", "Br", "I"), collapse = "|"))) %>% 
                        pull(CAS.Number)
```

Merging and categorizing all data
```{r}
# joining in the compound info (Inorganic, organic, halogenated. etc)
Envirotox_Comp_info_2 <-  Envirotox_Comp_info %>% 
  mutate(Substance_type = case_when(
    CAS.Number %in% Inorganic_vector_x ~ "Inorganic",
    grepl("mixt.", PesticideAI_name) ~ "Chemical mixture",
    is.na(CanonicalSMILES) ~ "Unknown",
      TRUE ~ "Organic"),
    `Heavy Metals` = case_when(CAS.Number %in% HM_vector_x ~ "1", 
      TRUE ~ "0"),
    Halogenated = case_when(CAS.Number %in% Halogenated_vector_x ~ "1", 
      TRUE ~ "0" ))

# # Defining a function to clean up the strings united as "Subgroup"
# clean_string <- function(x) {
#   x %>% 
#     str_split("; ") %>% 
#     unlist() %>% 
#     unique() %>% 
#     str_c(collapse = "; ")
# }
# Making a substance definition function to count the number of occurrences of a chemical use-category and extracting the top candidate.
# the function is named "use_prio_function()"
source("../code/Use_category_function.R")

# Running the function takes a couple of minutes.
#prio_use_cat <- use_prio_function(Envirotox_Comp_info_2)

Envirotox_Comp_info_3 <- cbind(Envirotox_Comp_info_2, 
                            prio_use_cat %>% 
                              mutate(
                                Select_use = as.character(Select_use),
                                n = as.numeric(n))
                            )

Envirotox_chem_prop_list_full <- Envirotox_Comp_info_3 %>% 
    # mutate_all( ~ na_if(., '')) %>% 
    # mutate_all( ~ na_if(., 'NA')) %>% 
    unite("Subgroup", c(
    "ChEBI_DB", "ATC_Type", "USEPA_ecotox_group", 
    "BCPC_activity", "DRUGBANK", "EPAPCS", "HEALTHY_BUILDING_NETWORK", 
    "NORMAN_ITN", "OPPIN", "PPDB", "USEPA", "Wikipedia"), 
    sep = "; ", na.rm = T, remove = F) %>%
  mutate(
    Group = case_when(
      grepl("pesticide", Select_use) ~ "Pesticide",
      grepl("antibiotic", Select_use) ~ "Antibiotic",
      grepl("PPCP", USEPA_ecotox_group, fixed = F) ~ "PPCP",
      grepl("Antiviral", Subgroup) ~ "Antiviral",
      grepl("Pharmaceutical", Subgroup) ~ "Pharmaceutical",
      #grepl("Pesticide", Subgroup) ~ "Pesticide",
      #grepl(paste(c("Antimicrobial", "antiinfective","antibacterial", "antibiotic"), collapse = "|"), Subgroup, fixed = F) ~ "Antibiotic",
        TRUE ~ case_when(
          Substance_type == "Organic" ~ "Other organic chemicals",
          is.na(CanonicalSMILES) ~ "Unknown",
            TRUE ~ "Other inorganic chemicals"))
    ) %>%
      # rowwise() %>% # Removing the Group category from Subgroup.
      # mutate(Subgroup = case_when(grepl(paste((Group)), Subgroup) ~ gsub(paste(Group, ";", sep=""), "", Subgroup),
      #                             TRUE ~ Subgroup)) %>% 
      # mutate(Subgroup = case_when(grepl(paste((Group)), Subgroup) ~ gsub(paste(Group), "", Subgroup),
      #                             TRUE ~ Subgroup)) %>% 
      # mutate(Subgroup = clean_string(Subgroup)) %>% 
      # ungroup() %>% 
      # mutate(
      #   Subgroup = gsub("_ingredient", "pesticide_ingredient", Subgroup),
      #   Subgroup = sub(";\\s*;", ";", Subgroup),
      #   Subgroup = sub("\\s*\\(s\\)", "", Subgroup),
      #   Subgroup = sub("\\(\\)", "", Subgroup),
      #   Subgroup = sub("^;+", "", Subgroup),
      #   Subgroup = sub("^ |^; |^;|^,| $|; *$", "", Subgroup)
      # ) %>% 
      select(CAS.Number, PesticideAI_name, CanonicalSMILES, Group, Subgroup, Select_use, n, Substance_type, `Heavy Metals`, Halogenated, 
             definition, ChEBI_DB, ATC_Type, USEPA_ecotox_group, BCPC_pesticide, BCPC_activity, ATC_Description,
             DRUGBANK, EPAPCS, HEALTHY_BUILDING_NETWORK, NORMAN_ITN, OPPIN, PPDB, USEPA, Wikipedia) %>% 
      mutate(Subgroup = tolower(Subgroup)) %>% 
      distinct(CAS.Number, .keep_all = T)
   
```


## PHYSICOCHEMICAL Properties 
### Wrangling operation of the QSAR Toolbox output
Close to all of the physicochemical information is queried from OECD QSAR Toolbox. The task to collect physicochemcal information for substances in the Envirotox database.
Since we dont need the overlapping substances, I first filter out the substances already occurring in PesticideAI.csv
**ONLY KEEPING mono-constituents**
```{r}
# Physchem wrangle function
source("../code/Physchem_read_wrangle_function.R")
# Processing and wrangling the raw data through function loaded above
# Reading and wrangling the query from OECD QSAR Toolbox
# ONLY KEEPING mono-constituents
Envirotox_physchem <- Physchem_read_wrangle_function(read.csv("../data/QSAR_Toolbox_physchem_data_Envirotox_2023-04-19_RAW.csv",header = T, sep = "\t", na.strings = "", fileEncoding = "UTF-16LE",  stringsAsFactors = FALSE))

# Merging of the two datasets
Envirotox_physchem_pest <- left_join(
    x = Envirotox_chem_prop_list_full,
    y = Envirotox_physchem,
    by = "CAS.Number") %>%
  select(CAS.Number, CanonicalSMILES, PesticideAI_name, Molecular.formula, Predefined.substance.type, 
         MW.g.mol, pKaChemClass, pKa.gain, pKa.loss, Kow_L.L, Koc_L.kg_MCI, kH25C_Pa.m3.mol, 
         Vapor.Pressure_Pa, Sol_mg.L, KdegA, KdegW,KdegSl, KdegSd, BAF_L.Kg, Source_Kow, 
         Source_Pvap, Source_Sol, Source_KH25C, Group, Subgroup, Select_use, n, Substance_type, `Heavy Metals`, Halogenated, definition)
 write.csv(Envirotox_physchem_pest, "../data/excel_references/Envirotox_physchem.csv", row.names = F)
```


